Multidimensional Persistence Module Classification via Lattice-Theoretic Convolutions

### 2020-10-09 paper/paper.tex
add some citations and more content to the paper

% ready for submission
% \usepackage{neurips_2020}

@@

Persistent homology has the ability to discern both the global topology \cite{}
and local geometry \cite{} of finite metric spaces (e.g. embedded weighted
graphs, point clouds in $\R^d$) making it a befitting feature for the purposes of
training a neural network. Single-dimensional homological persistence has drawn
recent attention in deep learning \cite{}. This is, in part, due to a wide range
of efficient software libraries [?,?,?] for computing barcodes (in the
``west-coast'' lingo) of filtrations (e.g. Rips, alpha, {C}ech, sub-levelset).
Barcodes provide a compact shape descriptor for metric space data \cite{} that
is stable with respect to Gromov-Hausdorff distance \cite{}. In a seminal paper
\cite{}, Carlsson and Zomorodian show there is no such compact description for
multi-parameter persistence.

\section{Backgound}
Due to space constraints, we can offer only a brief overview of multiparameter
persistent homology. For a primer on persistent homology,
see~\cite{ghrist_barcodes_2008,carlsson_topology_2009}; for multiparameter
persistent homology, see~\cite{}. An introduction to lattices may be found in~\cite{}.

@@

From $M$, we extract two types of features:
 the \introduce{Hilbert function}
  \[\text{Hilb}: L \to \Z_+;\quad \mathbf{x} \mapsto \dim(M_\mathbf{x}),\] and
the \introduce{multi-graded Betti numbers}\footnote{\textit{Caveat lector:} the
multi-graded Betti numbers are not the same as the topological Betti numbers
$\beta_i(\mathbb{X}) = \dim(H_i(\mathbb{X}))$.} $\xi_j: L \to \Z_+$, for $j =
0,1,2$. For $M = PH_i$ as above, the Hilbert function counts the number of
connected components ($i=0$), cycles ($i=1$), or higher dimensional voids ($i >
1$) of the complex $\mathbb{X}_{r,t}$ at each $(r,t) \in L$. The multi-graded
Betti numbers, on the other hand, capture information about locations of births
and deaths of persistence classes.

\subsection{Lattice-theoretic signal processing}

\section{Lattice Convolutional Neural Networks}

@@

To this end, we construct a lattice convolution-based neural network layer
suitable for use with features originating from multidimensional persistence
modules.

@@

@@

\section{Discussion}

\bibliographystyle{alpha}
\bibliography{sheafspectra}

>>>>

% ready for submission
%\usepackage{neurips_2020_tda}

@@

Persistent homology has the ability to discern both the global topology \cite{}
and local geometry \cite{} of finite metric spaces (e.g. embedded weighted
graphs, point clouds in $\R^d$) making it a befitting feature for the purposes of
training a neural network. Single-dimensional homological persistence has drawn
recent attention in deep learning \cite{hofer_deep_2017,pun_persistent-homology-based_2018,bruel-gabrielsson_topology_2020}. This is, in part, due to a wide range
of efficient software libraries \cite{otter_roadmap_2017,henselman_matroid_2017,bauer_ripser:_2019} for computing barcodes (in the
``west-coast'' lingo) of filtrations (e.g. Rips, alpha, {C}ech, sub-levelset).
Barcodes provide a compact shape descriptor for metric space data that
is stable with respect to Gromov-Hausdorff distance \cite{}. In a seminal paper
\cite{carlsson_theory_2009}, Carlsson and Zomorodian show there is no such compact description for
multi-parameter persistence.

\section{Backgound}
Due to space constraints, we can offer only a brief overview of multiparameter
persistent homology. For a primer on persistent homology,
see~\cite{ghrist_barcodes:_2008,carlsson_topology_2009}; for multiparameter
persistent homology, see~\cite{}. An introduction to lattices may be found in~\cite{}.

@@

While there does not exist a discrete set of invariants for $M$, we can extract
meaningful features. Two particularly informative types of features are 
 the \introduce{Hilbert function}
  \[\text{Hilb}: L \to \Z_+;\quad \mathbf{x} \mapsto \dim(M_\mathbf{x}),\] and
the \introduce{multi-graded Betti numbers}\footnote{\textit{Caveat lector:} the
multi-graded Betti numbers are not the same as the topological Betti numbers
$\beta_i(\mathbb{X}) = \dim(H_i(\mathbb{X}))$.} $\xi_j: L \to \Z_+$, for $j =
0,1,2$. For $M = PH_i$ as above, the Hilbert function counts the number of
connected components ($i=0$), cycles ($i=1$), or higher dimensional voids ($i >
1$) of the complex $\mathbb{X}_{r,t}$ at each $(r,t) \in L$. The multi-graded
Betti numbers, on the other hand, capture information about locations of births
and deaths of persistence classes.

\subsection{Lattice-theoretic signal processing}\label{sec:latticeconv}

\section{Lattice Convolutional Neural Networks}\label{sec:latticeCNN}

@@

To this end, we construct a lattice convolution-based neural network layer
suitable for use with features originating from multidimensional persistence
modules. We specialize the convolutions described in
Section~\ref{sec:latticeconv} to the particular case of regular finite
sublattices of $\R^2$. The meet and join operations are easily computed
elementwise:
\[(r,t) \wedge (r',t') = (\min(r,r'),\min(t,t'));\quad (r,t)\vee (r',t') =
  (\max(r,r'),\max(t,t')).\]

A lattice convolution layer takes as input an $N_{\text{in}}$-dimensional signal
$f : [m] \times [n] \to \R^{N_{\text{in}}}$ and outputs an
$N_{\text{out}}$-dimensional signal $ [m] \times [n] \to
\R^{N_{\text{out}}}$. The layer's parameters are given by a function $g: [m]
\times [n] \to \R^{N_{\text{out}}\times N_{\text{in}}}$. If we label the
entries of $f(x,y)$ by $f_i$ and the entries of $g(x,y)$ by $g^i_j$ The layer then acts by
\[\text{MeetConv}(f)(x,y)^j = \sum_{i} (f_i \ast_{\wedge} g^i_j)(x,y) = \sum_i
  \sum_{(a,b) \in [m]\times [n]} f_i(x \wedge a, y \wedge b)g^i_j(a,b)\]
in the case of convolution with respect to the meet operation, and
\[\text{JoinConv}(f)(x,y)^j = \sum_{i} (f_i \ast_{\vee} g^i_j)(x,y) = \sum_i
  \sum_{(a,b) \in [m]\times [n]} f_i(x \vee a, y \vee b)g^i_j(a,b)\]
in the case of convolution with respect to the join operation.

Convolutional neural networks are uesful in part because the convolution kernels
(here the functions $g$) can have very small support, enforcing locality and
reducing the number of parameters that must be learned. In the standard
convolutional setting, these kernels are implicitly supported in a neighborhood
of the origin, but the location of the kernel is not usually explicitly
specified. In the lattice setting, we do need to specify where the kernel
resides. Just as when we treat our domain as an abelian group, the kernel should
be supported near the identity, when we treat our domain as a lattice, the
kernel should be supported near the neutral element of the operation. That is,
for a meet convolution, $g$ should have support near the maximum $(m,n)$, and
for a join convolution, $g$ should have support near the minimum $(0,0)$.

@@

The pipline thus begins with a 3d model, which is sampled to produce a point
cloud in $\R^3$. This point cloud then produces a bifiltered simplicial complex,
whose persistent homology we calculate using RIVET~\cite{lesnick_interactive_2015}, producing
lattice-indexed signals given by the Hilbert function and the multi-graded Betti
numbers. These are then passed to the classifier, which produces a class prediction.

@@

We compare the performance of two convolutional networks on this classification
task. One uses the lattice-convolution based layers described in
Section~\ref{sec:latticeCNN}, and the other uses standard convolutional layers.
Each has three convolutional layers followed by three fully connected layers.

\section{Discussion}
Discuss.

\bibliographystyle{plain}
\bibliography{latticenn}
